Hypertension and microbiome

### 2019-06-12 articletwo.Rmd
Minor

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(Maaslin2)
library(pwr)
```

@@

    maaslin <- maaslinwrapper(pseq = pseq.genus.core.pruned,
                                    looped = var.BP,
                                    forced = var.CL,
                              taxa = taxa(pseq.genus.core.pruned))

@@

diversities <- mergediversities(alphadiversity, adonis.species)
```

@@

pcoa.df <- merge(pcoa.vectors %>% select(Axis.1, Axis.2, Axis.3),
                 meta(pseq.species), by=0, all = TRUE) %>%
    mutate(STATETREATMENT = factor(paste0(HYPERTENSION, ANYDRUG),
                                   levels = c("11", "10", "01", "00")),
           gMAP = cut(oMAP, seq(60, 160, 15)))

axis_labeller <- function(variable, value){
    axis.labels <- list("Axis.2" = sprintf("A. PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
                        "Axis.3" = sprintf("B. PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))
    return(axis.labels[value])
}

pcoa.plot <- ggplot(data=pcoa.df %>% gather(key, Axis, Axis.2, Axis.3),
                    aes(x=Axis.1, y=Axis, color = HYPERTENSION)) +
    facet_wrap(~key, labeller = axis_labeller) +
    geom_point(size=0.1, alpha=0.4) +
    xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
    ylab(sprintf("Subsequent PCoA Axes", 100*eig.fracs[3])) +
    scale_colour_manual(name = "State and treatment",
                        labels = c("Normotensive", "Hypertensive"),
                        breaks=c(0, 1),
                        values = c("blue", "red")) +   
    coord_fixed() +
    scale_x_continuous(breaks=seq(-1, 1, .25)) +
    scale_y_continuous(breaks=seq(-1, 1, .25)) +
    theme_classic() +
    guides(colour = guide_legend(override.aes = list(size=4))) +
    theme(text = element_text(size=14),
          legend.position = "none",
          strip.background = element_blank(),
          strip.placement = "outside")

ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=4, width=6, units = "in", dpi = 300)
```

@@

```{r pcoa axis analysis axis 1}
glm.pcoa <- lapply(c2l("Axis.1", "Axis.2", "Axis.3"), function(axis)
    calculateglm(pcoa.df,
                 modelstring = sprintf("%%s ~ %s + %s", axis, paste(var.CL, collapse = "+")),
                 filterstr = "Axis"))

do.call(rbind, glm.pcoa) %>% 
    select(response, term, estimate, std.error, p.value) %>%
    mutate_if(is.numeric, round, 3) %>%
    filter(p.value < 0.05) %>%
    arrange(p.value) %>%
    kable
```

@@

```{r maaslin}
maas.results.df.signf <- maaslin %>%
    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
                  Feature = gsub("_Bacteria", "", Feature),
                  beta = Coefficient) %>%
    dplyr::filter(Qval < 0.05) %>%
    arrange(Feature)

my.plot.order <- function(x) {
    y <- seq(length(x))
    y[x == "Systolic BP"] = 4
    y[x == "Diastolic BP"] = 3
    y[x == "Pulse pressure"] = 2
    y[x == "Mean arterial pressure"] = 1
    y[x == "Hypertension"] = 0
    y
}

maas.all.pairs <- expand.grid(Feature = maas.results.df.signf %>% pull(Feature),
                         Name = maas.results.df.signf %>% pull(Name))
maas.tables.df <- merge(maas.results.df.signf,
                        maas.all.pairs,
                        all.y=TRUE,
                        by.y=c('Feature', 'Name'),
                        by.x=c('Feature', 'Name')) %>%
    mutate(p.groups = cut(P.value*ifelse(Coefficient < 0, -1, 1),
                          breaks = c(-0.05, -0.01, -0.001, 0, 0.001, 0.01, 0.05),
                          labels = c("beta < 0, p < 0.05",
                                     "beta < 0, p < 0.01",
                                     "beta < 0, p < 0.001",
                                     "beta > 0, p < 0.001",
                                     "beta > 0, p < 0.01",
                                     "beta > 0, p < 0.05")))

@@

ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 3, width = 6, dpi = 300)
```

@@

@@

```{r forest plot lacto}
odds.hypertension.lactogroup
gsalt <- grid.arrange(
    myforestplot(adjmean.lacto.saltgroup,
                 xlab = "Dietary salt",
                 ylab = "Lactobacillus abundance",
                 subtitle = "A.",
                 my_y_scale = scale_y_continuous(limits = c(4.1,4.7), breaks = seq(4, 5, 0.1))),
    myforestplot(odds.hypertension.lactogroup,
                 xlab = "Lactobacillus abundance",
                 ylab = "Odds of hypertension",
                 subtitle = "B.",
                 my_y_scale = scale_y_log10(limits = c(0.4, 1.5), breaks = seq(0.5, 2.0, 0.5)),
                 interceptone = TRUE),
    ncol=2)

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(gridExtra)
library(png)
library(pander)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(pwr)
library(gtable)
```

@@

    maaslin <- maaslinwrapper(pseq = pseq.genus.core.pruned,
                                    looped = var.BP,
                                    forced = var.CL.min,
                              taxa = taxa(pseq.genus.core.pruned))

@@

diversities <- mergediversities(alphadiversity, adonis.species, names.dset = names.dset)
```

@@

pcoa.df <- merge(pcoa.vectors %>%
                 select(Axis.1, Axis.2, Axis.3),
                 meta(pseq.species), by=0, all = TRUE) %>%
    mutate(STATETREATMENT = factor(paste0(HYPERTENSION, ANYDRUG),
                                   levels = c("11", "10", "01", "00")),
           gMAP = cut(oMAP, seq(60, 160, 15))) %>%
    mutate_at(vars(starts_with("Axis.")), .funs = funs(myscale(.)))

axis_labeller <- function(variable, value){
    axis.labels <- list("Axis.2" = sprintf("PCoA Axis 2 (%.1f%%)", 100*eig.fracs[2]),
                        "Axis.3" = sprintf("PCoA Axis 3 (%.1f%%)", 100*eig.fracs[3]))
    return(axis.labels[value])
}

plot_labeller <-  function(value) {
        plot.labels <- list("Axis.2" = sprintf("A."),
                            "Axis.3" = sprintf("B."))
        return(plot.labels[value])
}

pcoa.df %>% pull(Axis.1) %>% summary

pcoa.plot <- ggplot(data=pcoa.df %>% gather(key, Axis, Axis.2, Axis.3),
                    aes(x=Axis.1, y=Axis, color = HYPERTENSION)) +
    facet_wrap(~key, scales = "free_y", strip.position = "left", labeller = axis_labeller) +
    geom_point(size=0.1, alpha=0.4) +
    xlab(sprintf("PCoA Axis 1 (%.1f%%)", 100*eig.fracs[1])) +
    ylab(NULL) +
    scale_colour_manual(name = "State and treatment",
                        labels = c("Normotensive", "Hypertensive"),
                        breaks=c(0, 1),
                        values = c("blue", "red")) +   
    scale_x_continuous(breaks=seq(-5, 5, 1)) +
    scale_y_continuous(breaks=seq(-5, 5, 1)) +
    theme_classic() +
    guides(colour = guide_legend(override.aes = list(size=4))) +
    coord_cartesian(clip = 'off') +
    geom_text(aes(x = -Inf, y = Inf, label = plot_labeller(key)),
              size = 5,
              hjust = 2,
              vjust = -1,
              inherit.aes = FALSE) +
    theme(text = element_text(size=12),
          legend.position = "none",
          strip.background = element_blank(),
          strip.placement = "outside",
          aspect.ratio = 1,
          strip.text = element_text(size = 14),
          axis.title.x = element_text(size = 14))

ggsave(file = "cache/pcoa-species.png", plot=pcoa.plot, height=3.4, width=6, units = "in", dpi = 300)
```

@@

```{r pcoa axis analysis axis 1}
glm.pcoa <- lapply(c2l("Axis.1", "Axis.2", "Axis.3"), function(axis)
    calculateglm(pcoa.df,
                 modelstring = sprintf("%%s ~ %s + %s", axis, paste(var.CL, collapse = "+")),
                 responses = list(model_1 = "oMAP", model_2 = "oSYSTM", model_3 = "oDIASM",
                                          model_4 = "oPULSEPRESSURE", model_5 = "HYPERTENSION"),
                 filterstr = "Axis"))

do.call(rbind, glm.pcoa) %>% 
    select(response, term, estimate, std.error, p.value, conf.low, conf.high) %>%
    mutate_if(is.numeric, round, 3) %>%
    filter(p.value < 0.05) %>%
    arrange(p.value) %>%
    kable

@@

```{r maaslin clean results}
maas.results.df.signf <- maaslin %>%
    merge(names.dset %>% select(Covariate, Category, Name), by.x ="Variable", by.y = "Covariate") %>%
    dplyr::mutate(Qval = p.adjust(P.value, method="BH"),
                  Feature = gsub("[gs]_", "", Feature),
                  beta = Coefficient) %>%
    dplyr::filter(Qval < 0.05) %>%
    arrange(Feature)
```

```{r maaslin plot}
my.plot.order <- function(x) {
    y <- seq(length(x))
    y[x == "Systolic BP"] = 4
    y[x == "Diastolic BP"] = 3
    y[x == "Pulse pressure"] = 2
    y[x == "Mean arterial pressure"] = 1
    y[x == "Hypertension"] = 0
    y
}

maas.all.pairs <- expand.grid(Feature = maas.results.df.signf %>% pull(Feature),
                         Name = maas.results.df.signf %>% pull(Name))

maas.tables.df <- merge(maas.results.df.signf,
                        maas.all.pairs,
                        all.y=TRUE,
                        by.y=c('Feature', 'Name'),
                        by.x=c('Feature', 'Name')) %>%
    mutate(p.groups = cut(P.value*ifelse(Coefficient < 0, -1, 1),
                          breaks = c(-0.05, -0.01, -0.001, 0, 0.001, 0.01, 0.05),
                          labels = c("beta < 0, p < 0.05",
                                     "beta < 0, p < 0.01",
                                     "beta < 0, p < 0.001",
                                     "beta > 0, p < 0.001",
                                     "beta > 0, p < 0.01",
                                     "beta > 0, p < 0.05")))

@@

ggsave(file = "cache/maaslin.png", plot = maaslinplot, height = 30, width = 40, unit = "in", dpi = 300)
```

@@

```{r multi model}
abund.genus.systm <- pseq.genus %>%
    abundances(., transform = "compositional") %>%
    t %>%
    as.data.frame %>%
    tibble::rownames_to_column("id")

df.genus.systm <- meta(pseq.genus.systm) %>%
    tibble::rownames_to_column("id") %>%
    full_join(., abund.genus.systm, by = "id") %>%
    mutate(Anaerotruncus = scale(asin(sqrt(g_Anaerotruncus))),
           Sutterella = scale(asin(sqrt(g_Sutterella))))

fo.genus.systm <- sprintf("oSYSTM ~ Anaerotruncus + Sutterella + Anaerotruncus:Sutterella + %s",
                          paste0(var.CL, collapse = " + "))

lm(as.formula(fo.genus.systm), data = df.genus.systm) %>% summary

df.genus.systm %>%
    mutate(diff = -0.75418 * Anaerotruncus + 0.89265 * Sutterella) %>%
    pull(diff) %>%
    summary

@@

```{r forest lims, eval = FALSE, include = FALSE}
adjmean.lacto.saltgroup
```

```{r forest plot lacto}
odds.hypertension.lactogroup
gsalt <- grid.arrange(
    myforestplot(adjmean.lacto.saltgroup,
                 xlab = "Dietary salt",
                 ylab = "Lactobacillus abundance",
                 subtitle = "A.",
                 my_y_scale = scale_y_continuous(limits = c(1.15,1.42), breaks = seq(0, 5, 0.1))),
    myforestplot(odds.hypertension.lactogroup,
                 xlab = "Lactobacillus abundance",
                 ylab = "Odds of hypertension",
                 subtitle = "B.",
                 my_y_scale = scale_y_log10(limits = c(0.4, 1.5), breaks = seq(0.5, 2.0, 0.5)),
                 interceptone = TRUE),
    ncol=2)
